\begin{thebibliography}{10}
\addtolength{\leftmargin}{0.2in} % sets up alignment with the following line.
\setlength{\itemindent}{-0.2in}

\bibitem{FWJA} Atsushi Igarashi, Benjamin C. Pierce, and Philip Wadler. Featherweight java: a minimal core calculus for java and GJ. {\it ACM Trans. Program. Lang. Syst., 23(3):396–450}, 2001. {\tt URL: http://doi.acm.org/10.1145/503502. 503505, doi:10.1145/503502.503505}.
\bibitem{FWJS} Thomas H. Austin, Tim Disney, Alan Jeffrey, and Cormac Flanagan. Dynamic information flow analysis for featherweight javascript, 2011.
\bibitem{EEL} Roberto Ierusalimschy, Luiz Henrique de Figueiredo, and Waldemar Celes Filho. Lua an extensible extension language. {\it Softw., Pract. Ex- per., 26(6):635–652}, 1996. {\tt URL: http://dx.doi.org/10.1002/(SICI) 1097-024X(199606)26:6<635::AID-SPE26>3.0.CO;2-P}.
\bibitem{LRM} Lua 5.2 Reference Manual, {\it Roberto Ierusalimschy, Luiz Henrique de Figueiredo, Waldemar Celes}, 2013.
\bibitem{EOL} Roberto Ierusalimschy, Luiz Henrique de Figueiredo, and Waldemar Celes. The evolution of an extension language: A history of lua, {\it In Proceeding of V Brazilian Symposium on Programming Language, pages 14–28}, 2001.
\bibitem{TEL} Roberto Ierusalimschy, Luiz Henrique de Figueiredo, and Waldemar Celes Filho. The evolution of lua. {\it In Proceedings of the Third ACM SIGPLAN History of Programming Languages Conference (HOPL-III), San Diego, California, USA, 9-10 June 2007, pages 1–26}, 2007. {\tt URL: http://doi.acm. org/10.1145/1238844.1238846, doi:10.1145/1238844.1238846}.
\bibitem{AIL} Matheson, Ash, An Introduction to Lua. {\it GameDev.net. Accesat la 3 ianuarie 2013}, 2003.
\bibitem{WLA} Paul Emmerich, Beginning Lua with World of Warcraft Addons, {\it ISBN: 9781430223719}, 2009.
\bibitem{WCU} Curse Inc. Curse.com, Recently accessed in 2015. {\tt URL: http://www.curse.com/
addons/wow}.
\bibitem{CLFL} CloudFlare, Lua: the world's most infuriating language, 2013.
\bibitem{APLL} Adobe Photoshop Lightroom \& Lua, 2007, Recently accessed in 2015. {\tt URL: http://blog.kodekabuki.com/post/29578340/adobe-photoshop-lightroom
-lua}.
\bibitem{NPL} Pushing Nginx to its limit with Lua, 2012. {\tt URL: http://blog.cloudflare
.com/pushing-nginx-to-its-limit-with-lua/, Matthieu Tourne}.

\bibitem{LJT} Mike Pall, Introduction to LuaJIT, Recently accessed in 2015. {\tt URL: http://
wiki.luajit.org/New-Garbage-Collector}.
\bibitem{FCV} Well House Consultants LTD. Functions are first class variables in Lua and Python, Recently accessed in 2015. 
{\tt URL: http://www.wellho.net/mouth/3695
\_Functions-are-first-class-variables-in-Lua-and-Python.html}.
\bibitem{EOJS} Arjun Guha, Claudiu Saftoiu, and Shriram Krishnamurthi. The essence of javascript. {\it In ECOOP 2010 - Object-Oriented Programming, 24th European Conference, Maribor, Slovenia, June 21-25, 2010. Proceedings, pages 126– 150}, 2010. {\tt URL: http://dx.doi.org/10.1007/978-3-642-14107-2\_7, doi:10.1007/978-3-642-14107-2\_7}.
\bibitem{AIOS} Hayo Thielecke, University of Birmingham, An introduction to operational semantics and abstract machines, 2012.
\bibitem{CISCO} Cisco Inc, Cisco ASA Software SharePoint RAMFS Integrity and Lua Injection Vulnerability, 2014.
\bibitem{PIL} Roberto Ierusalimschy, Programming in Lua, Third Edition,  2013.
\bibitem{DA} Mozilla Developer Network, Destructuring assignment, Recently accessed in 2015. {\tt URL: https://developer.mozilla.org/en-US/docs/Web/JavaScript/
Reference/Operators/Destructuring\_assignment}.
\bibitem{LC} Raul Rojas, A Tutorial Introduction to the Lambda Calculus, 1997.
\bibitem{TAPL} Benjamin C. Pierce. Types and programming languages: The next generation. {\it In 18th IEEE Symposium on Logic in Computer Science (LICS 2003), 22-25 June 2003, Ottawa, Canada, Proceedings, page 32}, 2003. {\tt URL: http://dx.doi.org/10.1109/LICS.2003.1210042, doi:10.1109/ LICS.2003.1210042}.
\bibitem{LTC} Lua Type Checking, Recently accessed in 2015. {\tt URL: http://lua-users.org/
wiki/LuaTypeChecking}.
\bibitem{LuaScoping} Lua Scoping, Recently accessed in 2015. {\tt URL: http://lua-users.org/wiki/
LuaScoping}.
\bibitem{begLua} Kurt Jung, Aaron Brown, Beginning Lua Programming, 2012.
\bibitem{PIJS} Douglas Crockford, Prototypal Inheritance in JavaScript, 2008. {\tt URL: http://
javascript.crockford.com/prototypal.html}.
\bibitem{PIH} Graham Hutton, Programming in Haskell, 2007.
\bibitem{WCAI} Ronald Mak, Writing Compilers and Interpreters: A Software Engineering Approach, 2009.
\bibitem{looseC} Doug Kaye, Loosely Coupled: The Missing Pieces of Web Services, 2008.
\bibitem{IHAS} Instruction of Haskell, 2008. {\tt URL: https://hackage.haskell.org/package/
parsec-3.1.0/docs/Text-ParserCombinators-Parsec.html}.
\\

\end{thebibliography}
